\acknowledgements
{
I am grateful for all the opportunities and resources I received at IIT-H. I am most grateful to my parents for providing me with academic freedom and support. I am grateful to Prof. Vineeth N. Balasubramanian for initiating my research journey and for always being there for students. I would like to thank Prof. J. Saketha Nath, my PhD adviser, for his courses where I learned the basics and for his guidance in my projects, especially in Submodular-Attribution (which forced me to forgo my biases) and MMD-OT (which shaped my research outlook). I always longed for someone to review my proof attempts; I thank Sir for giving so much time to these. Through Sir, I learned things, in and beyond academics, that might have taken me ages otherwise.

I am grateful to my Doctoral Committee for their constructive feedback and especially to Prof. M. Vidyasagar for inspiring us with his commitment to research. I am grateful to Prof. P.N. Karthik and Prof. Ganesh Ghalme for sharing their expertise and being so approachable. I also thank Google for my fellowship. I would like to thank Dr. Pratik Jawanpuria for helping me scale some of our Algos and for Chap.\ref{chap:ch2}'s idea. 
For our discussions, I thank  Siddharth Shrivastava, A.V. Mahvith Kusumakar, Vishwak Srinivasan, Harshitha Machiraju, Sai Vikas Desai, Aditya Chattopadhyay, Sai Srinivas Kancheti, Dr. A. Ravi Sankar, Radhika Dua, Shivam Chandhok, Pranav K. Nayak, Gautham Bellamkonda, Kartik Srinivas, Dr. A. Gowtham Reddy, S. Venkatakeerthy, Dr. Maruthi S. Inukonda, my 12 y/o cousin and the COT-coauthors.
I am grateful to the Late Mr. Ajay Singh for his narratives of those scientific discoveries that made me dream of a research life in '13. I thank Mr. Pradeep Besara, Prof. Kuldeep Sharma, Mr. Abhyudaya A. Mishra and Mr. Prashant Agarwal for their guidance during different stages, and my friends Sana \& Subhra for their company.
Finally, I express my deepest gratitude to those who extended their support after I had given up on everything following the sudden demise of my father. After returning back, I spent weeks in despair before Dr. Suvodip Dey came as the first one to guide me, which I will never forget. I am grateful to Prof. Ramakrishna Upadrasta, Dr. Roopam Saxena and Dr. Manisha Dubey for being so kind and supportive, and to Sai Srinivas Kancheti for being there to hear me out. 
%The first thing that helped me was an impromptu conversation with HG Amogh Lila Prabhu. I thank IIT-H for that event in Jan '24. 
And, I am extremely grateful to Prof. Ganesh Ghalme for giving me the project that revived my academic interests.

My academic journey is due to my father, who never stopped believing in me. While he could not sense when I arrived, I will cherish our Acad-C-block calls, including our casual video call, interrupted by the rains, which turned out to be the last one.
I'll prepare to safeguard his readings better next time.
I place this thesis for my father's journey onward. This is also in the memory of my late grandparents.
}