\chapter*{List of Publications}
\phantomsection\addcontentsline{toc}{chapter}{List of Publications}
\vspace{-0.9in}
\small
\begin{enumerate}
\itemsep-0.5em
    \item  \href{https://openreview.net/forum?id=eN9CjU3h1b}{MMD-Regularized Unbalanced Optimal Transport} \textbf{(Chapter \ref{chap:ch1})}\newline
    \underline{P Manupriya}, SN Jagarlapudi, P Jawanpuria. \newline  \textit{Transactions on Machine Learning Research}, TMLR Journal Jan '24.
     \newline
     {\href{https://drive.google.com/file/d/1COByfGSeWplgM1J-Xu22bhXZYovxLfcc/view?usp=drive_link}{\color{black}{\textit{Poster}}} \textit{accepted at WiML Symposium, ICML '24.}}\\
      \href{https://github.com/Piyushi-0/MMD-reg-OT}{\color{darkgray!90}{Code - https://github.com/Piyushi-0/MMD-reg-OT}}
    \item \href{https://arxiv.org/pdf/2406.04914}{Submodular Framework for Structured-Sparse Optimal Transport} \textbf{(Chapter \ref{chap:ch2})} \newline
\underline{P Manupriya}, P Jawanpuria, KS Gurumoorthy, SN Jagarlapudi, B Mishra. \newline 
 \textit{International Conference on Machine Learning}, ICML '24.\newline \href{https://github.com/Piyushi-0/Sparse-UOT}{\color{darkgray!90}{Code - https://github.com/Piyushi-0/Sparse-UOT}}
    \item \href{https://arxiv.org/pdf/2305.15901}{Consistent Optimal Transport with Empirical Conditional Measures} \textbf{(Chapter \ref{chap:ch3})}\newline
    \underline{P Manupriya}, RK Das, S Biswas, SN Jagarlapudi. 
    \newline
     \textit{International Conference on Artificial Intelligence and Statistics}, AISTATS '24.
    \newline \href{https://github.com/atmlr-lab/COT}{\color{darkgray!90}{Code - https://github.com/atmlr-lab/COT}}
\end{enumerate}
\vspace{-0.05in}
\paragraph{Other manuscripts from the candidate's Master's-PhD dual degree:}
\vspace{-0.2in}
\begin{itemize}
\itemsep-0.5em
    \item Multi-agent Multi-armed Bandits with Minimum Reward Guarantee Fairness
    \newline\underline{P Manupriya}, Himanshu, SN Jagarlapudi, G Ghalme. 
    \newline \textit{(Under review at an $A^*$ conference).}
    \newline  \href{https://github.com/Piyushi-0/Fair-MAMAB}{\color{darkgray!90}{Code - https://github.com/Piyushi-0/Fair-MAMAB}} 
    \item  \href{https://arxiv.org/pdf/1902.02302}{Neural Network Attributions: A Causal Perspective}\newline
    A Chattopadhyay, \underline{P Manupriya}, A Sarkar, VN Balasubramanian.\newline
     \textit{International Conference on Machine Learning}, Oral at ICML '19.
    \newline \href{https://github.com/Piyushi-0/ACE}{\color{darkgray!90}{Code - https://github.com/Piyushi-0/ACE}}
    \item \href{https://proceedings.mlr.press/v151/manupriya22a.html}{Improving Attribution Methods by Learning Submodular Functions} \newline\underline{P Manupriya}, TR Menta, SN Jagarlapudi, VN Balasubramanian. \newline \textit{International Conference on Artificial Intelligence and Statistics}, AISTATS '22.
    \newline
     \href{https://github.com/Piyushi-0/SEA-NN}{\color{darkgray!90}{Code - https://github.com/Piyushi-0/SEA-NN}}\newline
     Precursor: \href{http://interpretable-ml.org/icml2020workshop/pdf/29.pdf}{SEA-NN: Submodular Ensembled Attribution for Neural Networks}
    \newline
     \underline{P Manupriya}, SN Jagarlapudi, VN Balasubramanian.
    \newline  \textit{ICML 2020 Interpretability workshops: WHI (Spotlight), XXAI (Poster).}
    \item ``Improving Fairness in Matching under Uncertainty", \href{https://neurips.cc/virtual/2024/101544}{Poster} for an extended abstract accepted at the \textit{AFME Workshop, NeurIPS '24}. 
    \newline
    \underline{P Manupriya}.
\end{itemize}